---
title: "**PBSR - Assignment 2**"
author: "Shubhangi Sanyal<br>MDS202238"
date: '2022-11-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 5: Computational Finance - Modelling Stock prices

Following piece of code download the prices of TCS since 2007

```{r}
library(quantmod)
getSymbols('TCS.NS')
tail(TCS.NS)
```
Plot the adjusted close prices of TCS
```{r}
plot(TCS.NS$TCS.NS.Adjusted)
```

**Download the data of market index Nifty50**. The Nifty 50 index indicates how the over all market has done over the similar period.
```{r}
getSymbols('^NSEI')
tail(NSEI)
```
Plot the adjusted close value of Nifty50
```{r}
plot(NSEI$NSEI.Adjusted)
```


### Log-Return 
We calculate the daily log-return, where log-return is defined as
$$
r_t=\log(P_t)-\log(P_{t-1})=\Delta \log(P_t),
$$
where $P_t$ is the closing price of the stock on $t^{th}$ day.

```{r}
TCS_rt = diff(log(TCS.NS$TCS.NS.Adjusted))
Nifty_rt = diff(log(NSEI$NSEI.Adjusted))
retrn = cbind.xts(TCS_rt,Nifty_rt) 
retrn = na.omit(data.frame(retrn))

plot(retrn$NSEI.Adjusted,retrn$TCS.NS.Adjusted
     ,pch=20
     ,xlab='Market Return'
     ,ylab='TCS Return'
     ,xlim=c(-0.18,0.18)
     ,ylim=c(-0.18,0.18))
grid(col='grey',lty=1)
```

+ Consider the following model:

$$
r_{t}^{TCS}=\alpha + \beta r_{t}^{Nifty} + \varepsilon,
$$
where $\mathbb{E}(\varepsilon)=0$ and $\mathbb{V}ar(\varepsilon)=\sigma^2$.

1. Estimate the parameters of the models $\theta=(\alpha,\beta,\sigma)$ using the method of moments type plug-in estimator discussed in the class.

### Answer:

For simplicity, the model equation has been written as:<br>
$y_t=\alpha+\beta x_t+\varepsilon$ <br>

where, $y_t = r_t^{TCS}$ and $x_t=r_t^{Nifty}$ 

For the method of moments, the following three moment conditions are followed in order to estimate the unknown parameters $\alpha$, $\beta$ and $\sigma$.<br>
 
**(1)** $\mathbb{E}(y_t-\alpha-\beta x_t)=0$ <br>

**(2)** $\mathbb{E}(x_t(y_t-\alpha-\beta x_t))=0$ <br>

**(3)** $\mathbb{E}(\varepsilon^2)=0$ <br>

Using the above moment conditions, we have the following forms of $\alpha$, $\beta$ and $\sigma$,<br>

$\hat{\alpha}=\bar{y_t}-\hat{\beta} \bar{x_t}$,<br>

$\hat{\beta} = \frac{\sum_{t=1}^{n}(x_t-\bar{x})(y_t-\bar{y})}{\sum_{t=1}^{n}(x_t-\bar{x})^2}$, and <br>

$\sigma^2 = \mathbb{V}ar(\varepsilon) = \mathbb{V}ar(y_t-\hat{\alpha}-\hat{\beta} x_t)$

```{r}
#MoM
x <- retrn$NSEI.Adjusted
y <- retrn$TCS.NS.Adjusted
n <- length(x)
betahat <- sum((x-mean(x))*(y-mean(y)))/sum((x-mean(x))**2)
alphahat <- mean(y)-betahat*mean(x)
yhat <- alphahat + betahat*x
resi <- y-yhat
sigma <- (sum(resi**2)/(n-1))**0.5
```


2. Estimate the parameters using the `lm` built-in function of `R`. Note that `lm` using the OLS method.

### Answer:

```{r}
#linear model using OLS
model<-lm(retrn$TCS.NS.Adjusted~retrn$NSEI.Adjusted,data=retrn)
summary(model)
```

3. Fill-up the following table

### Answer:

```{r}
#Filling table
MM_OLS <- data.frame(Parameters=c('alpha','beta','sigma'),
                     MoM=c(alphahat,betahat,sigma),
                     OLS=c(model$coefficients[1],model$coefficients[2],var(model$residuals)**0.5))
MM_OLS
```

Parameters |Method of Moments |OLS
-----------|------------------|-------------
   $\alpha$|      0.0004628227|0.0004628227
    $\beta$|      0.7436835788|0.7436835788
   $\sigma$|      0.0161868605|0.0161868605


4. If the current value of Nifty is 18000 and it goes up to 18200. The current value of TCS is Rs. 3200/-. How much you can expect TCS price to go up?

### Answer:

For this, the estimated parameters can be used to calculate the return of TCS which can give the expected price of TCS as follows,

$r_t^{Nifty}=\log(P_t^{Nifty})-\log(P_{t-1}^{Nifty})=\log(18200)-log(18000)$<br>

$r_t^{TCS}=\hat{\alpha}+\hat{\beta} r_t^{Nifty}$<br>

$P_t^{TCS}= e^{r_t^{TCS}+\log(P_{t-1}^{TCS})} = e^{r_t^{TCS}+\log(3200)}$<br>

```{r}
# Predicting TCS price
a <- log(18200)-log(18000)
tcs_hat<-alphahat+betahat*a
b<-exp(tcs_hat+log(3200))

cat("Thus, the expected TCS price is ", b)
```